---
title: "Regression_Case"
author: "Chun-Li Hou"
date: "2020/8/24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(dplyr, ggplot2, reshape)
```

## Introduction

We’ll look at a College data and my assignment is to study how income varies across college major categories.

A codebook for the dataset is given below:

- rank: Rank by median earnings
- major_code: Major code
- major: Major description
- major_category: Category of major
- total: Total number of people with major
- sample_size: Sample size of full-time, year-round individuals used for income/earnings estimates: p25th, median, p75th
- p25th: 25th percentile of earnings
- median: Median earnings of full-time, year-round workers
- p75th: 75th percentile of earnings
- perc_men: % men with major (out of total)
- perc_women: % women with major (out of total)
- perc_employed: % employed (out of total)
- perc_employed_fulltime: % employed 35 hours or more (out of employed)
- perc_employed_parttime: % employed less than 35 hours (out of employed)
- perc_employed_fulltime_yearround: % employed at least 50 weeks and at least 35 hours (out of employed and full-time)
- perc_unemployed: % unemployed (out of employed)
- perc_college_jobs: % with job requiring a college degree (out of employed)
- perc_non_college_jobs: % with job not requiring a college degree (out of employed)
- perc_low_wage_jobs: % in low-wage service jobs (out of total)

The specifically question for this project is: “Is there an association between college major category and income?”

Based on your analysis, would you conclude that there is a significant association between college major category and income?

## Environment setting

```{r}
# install.packages('devtools')
devtools::install_github('jhudsl/collegeIncome')
library(collegeIncome)
data(college)
glimpse(college)
```

## Analysis

```{r}
college$major = as.factor(college$major)
college$major_code = as.factor(college$major_code)
college$major_category = as.factor(college$major_category)
college = college %>% select(major_code, 
                             major_category, 
                             major, 
                             median)
glimpse(college)
head(college)
```

```{r, fig.height = 5, fig.width = 7}
boxplot(median/1000 ~ major_category, 
        data = college, 
        main = 'Income vs. Major', 
        ylab = 'Income (thousands of dollar)',
        xlab = '',
        las = 2)
```

We can see the distribution of the median of Income of each major is not normal, they’re skewed. However for the purpose of this project of practicing linear model, I assume they’re normal.

```{r}
unique(college$major_category)
college = college[order(college$major_category), ]
major_category_ref = relevel(college$major_category, "Arts")
fit = lm(median ~ major_category_ref, data = college)
summary(fit)$coef
```

```{r}
A = matrix(, nrow = 16, ncol = 16)

for (i in 1:16){
    major_category_ref = relevel(college$major_category, as.character(unique(college$major_category)[i]))
    fit = lm(median ~ major_category_ref, data = college)
    tmp = summary(fit)$coef[,4]
    # swap the first element to the corresponding position in the diagonal matrix
    tmp1 = tmp[1:i]
    tmp1 = c(0,tmp1)
    tmp1 = c(tmp1[-2],tmp1[2])
    tmp1 = tmp1[-1]
    # save to A
    A[,i] = c(tmp1,tmp[-(1:i)])
}

B = data.frame(A)
names(B) = unique(college$major_category)
B$major = unique(college$major_category)
Bmelt = melt(B)
```

```{r}
g = ggplot(data = Bmelt, aes(x = variable, y = major, fill = value))
g = g + geom_tile()
g = g + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylab("Major") + xlab("Major")
g = g + ggtitle("Probability of difference in Income between Majors")
g = g + coord_fixed()
g
```

```{r}
g = ggplot(data = Bmelt, aes(x = variable, y = major, fill = value < 0.025))
g = g + geom_tile()
g = g + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylab("Major") + xlab("Major")
g = g + ggtitle("Difference in Income between Majors")
g = g + coord_fixed()
g
```

```{r}
major_category_ref = relevel(college$major_category, "Business")
fit = lm(median ~ major_category_ref, data = college)
summary(fit)$coef
```

```{r}
business_diff = summary(fit)$coef[-1,]
business_diff[order(business_diff[,4])[1:5], 4]
```

